### 2018-11-14 Sections/network_algorithms.tex
Fig Dijkstra update

The first proven shortest-path algorithm was invented by and named after Dijkstra \cite{bib:Dijkstra59}, which requires runtime,
\begin{align}
	O(|V|^2),
\end{align}
which has since been improved to,
\begin{align}
	O(|E|+|V|\log |V|),	
\end{align}
by \cite{bib:FredmanLawrence84}, residing in \textbf{P}\index{P} (one of the relatively few, and highly valuable optimisation problems that is classically efficient). Subsequently, a number of improvements and variations on Dijkstra's algorithm have been proposed, most notably the $A^*$ algorithm \cite{bib:Astar}\index{Shortest-path algorithm}, which has found widespread modern use, using a heuristic approach to improve performance over Dijkstra. An alternate implementation of Dijkstra using

@@

\begin{figure}[!htbp]
\if 2\pubmode
\includegraphics[width=0.475\textwidth]{dijkstra_example_long}
\else
\includegraphics[width=0.7\textwidth]{dijkstra_example}
\fi
\caption{Example of Dijkstra's shortest-path algorithm for finding the shortest route from $A$ to $B$, yielding a shortest route of distance 20 using 5 algorithmic steps. Initially (\mbox{$t=0$}) all nodes are designated a tentative distance of $\infty$ and marked as unvisited (blue vertices), except the starting node $A$, which is designated distance $0$ and visited (red vertices). At each step we choose the lowest cost path to a previously unexplored node (red edge). This updates the tentative distance of the newly-visited neighbour, which is now marked as visited. This is iterated until all nodes have been visited. Each node is visited exactly once.}\label{fig:dijkstra_eg}\index{Shortest-path algorithm}	
\end{figure}

>>>>

The first proven shortest-path algorithm was invented by and named after Dijkstra \cite{bib:Dijkstra59}, which requires runtime,
\begin{align}
	O(|V|^2),
\end{align}
which has since been improved to,
\begin{align}
	O(|E|+|V|\log |V|),	
\end{align}
by \cite{bib:FredmanLawrence84} using min-priority queues\index{Min-priority queues} implemented via Fibonacci heaps\index{Fibonacci heaps}. Thus, the shortest-path algorithm resides in \textbf{P}\index{P} -- one of the relatively few, and highly valuable optimisation problems that is classically efficient. Subsequently, a number of improvements and variations on Dijkstra's algorithm have been proposed, most notably the $A^*$ algorithm \cite{bib:Astar}\index{Shortest-path algorithm}, which has found widespread modern use, using a heuristic approach to improve performance over Dijkstra. An alternate implementation of Dijkstra using

@@

\begin{figure}[!htbp]
\if 2\pubmode
\includegraphics[width=0.475\textwidth]{dijkstra_example_long}
\else
\includegraphics[width=0.7\textwidth]{dijkstra_example}
\fi
\caption{Example of Dijkstra's shortest-path algorithm for finding the shortest route from $A$ to $B$ (blue route), yielding a shortest route of distance 20 using 5 algorithmic steps. Initially (\mbox{$t=0$}) all nodes are designated a tentative distance of $\infty$ and marked as unvisited (blue vertices), except the starting node $A$, which is designated distance $0$ and visited (red vertices). At each step we choose the lowest cost path to a previously unexplored node (red edges). This updates the tentative distance of the newly-visited neighbour, which is now marked as visited. This is iterated until all nodes have been visited. Each node is visited exactly once.}\label{fig:dijkstra_eg}\index{Shortest-path algorithm}	
\end{figure}
